Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2024-12-20 analysis/paper/article.qmd
In which Jess experiences the horrors of Quarto and GitHub

<!-- Instructors to authors for AAQ

@@

-->

@@

The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States. We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.

@@

One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials so that the hiring committee can focus on evaluating and ranking the specific accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program. This has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit. For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans. In a comparison of job ads from 1999–2000 and 2019–2020 published in _Anthropology News_ @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants. They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy. In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements, requesting nine and thirteen documents from applicants. This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.

@@

Our primary data source is the Archaeology Academic Jobs Wiki. Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company. The Archaeology pages are part of the Academic Jobs Wiki, which coordinates similar collaboratively-edited resources for around 40 academic disciplines. The coordinators and contributors are nearly all anonymous or pseudonymous. Typically contributors copy and paste the text of job ads from other sources, such as the _Chronicle of Higher Education_, _Higher Ed Jobs_, and university websites, into the wiki, collecting ads originally posted in numerous different locations. Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc. Contributors also edit the page to ask and answer questions about the positions and the application process. These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market. Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g. @Passalacqua_2018; @musial2018five] and a widely recommended resource [e.g. @lightfoot2021preparing].

>>>>

```{=html}
<!-- Instructors to authors for AAQ

@@

-->
```

@@

The aim of this paper is to TEACH JESS HOW TO USE QUARTO AND GITHUB, UNFORTUNATELY FOR JESS. We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.

@@

One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials so that the hiring committee can focus on evaluating and ranking the specific accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program. This has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit. For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans. In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News* @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants. They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy. In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements, requesting nine and thirteen documents from applicants. This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.

@@

Our primary data source is the Archaeology Academic Jobs Wiki. Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company. The Archaeology pages are part of the Academic Jobs Wiki, which coordinates similar collaboratively-edited resources for around 40 academic disciplines. The coordinators and contributors are nearly all anonymous or pseudonymous. Typically contributors copy and paste the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting ads originally posted in numerous different locations. Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc. Contributors also edit the page to ask and answer questions about the positions and the application process. These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market. Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g. @Passalacqua_2018; @musial2018five] and a widely recommended resource [e.g. @lightfoot2021preparing].
